\LARGE \bf Energy-Aware Dynamic Planning Algorithm for Autonomous UAVs

### 2020-12-01 icra-2021.tex
working on the text (2)

\begin{figure*}[t]
  \centering
  \input{\figpath/icra-2021_overview2.tikz}
  \caption{The mission consists of circles and lines covering a polygon. The UAV heads to $\varphi_0$ initially, using the desired velocity vector $\dot{\mathbf{p}}_d$. It later heads to $\varphi_0^-$ reducing the radius $c_{0,1}$ to satisfy the battery constraints. The UAV then converges to $\varphi_1$ in stage $\mathcal{M}_1$, $\varphi_2$ in stage $\mathcal{M}_2$, and so on (the circle $\varphi_2$ is not visible in the figure; it connects $\varphi_1$ and $\varphi_3$).}
  \label{fig:overview}
\end{figure*}

@@

Many scenarios involving unmanned aerial vehicles (UAVs), such as precision agriculture, search and rescue, and surveillance, require high autonomy but have limited energy budgets. A typical example of these scenarios is a UAV flying a path and performing some on-board computational tasks. For instance, the UAV might detect ground patterns and inform other ground-based actors with little human interaction. We refer to the computational tasks that can be dynamically replanned as computations. We are interested in energy-aware dynamic planning, which would find optimal tradeoffs between the path, computations, and energy requirements. Current generic planning solutions for outdoor UAVs do not plan path and computations dynamically, nor are energy-aware. They are often semi-autonomous; the plan is static and usually defined using planning software~\cite{daponte2019review}. Such a state of practice has prompted us to propose an \emph{energy-aware dynamic planning algorithm} for UAVs. The algorithm combines and generalizes some of the past body of knowledge on mobile robot planning problems and addresses the increasing \emph{computational demands} and their relation to energy consumption, path, and autonomy for the UAV planning problem.

@@

Unlike most of the past planning algorithms literature, our algorithm plans both the path and computations. We guide the UAV using a vector field~\cite{de2017guidance} that converges smoothly to a planned trajectory. The use of vector fields for planning is widely discussed in the literature~\cite{lindemann2005smoothly,gonccalves2010vector,panagou2014motion,zhou2014vector,kapitanyuk2017guiding}. We denote the trajectory with a mathematical function that we refer to as the trajectory explicit equation (TEE). The path is composed of multiple TEEs. We express the computations with a modeling tool presented in our previous work~\cite{seewald2019coarse}. The algorithm alters the TEE and computations within some parameters using output model predictive control (MPC). These parameters are the maximum, minimum, and current levels of the TEE and computations. The user specifies the parameters in an initial plan with an energy budget. The energy budget is the battery capacity and some other battery parameters. In the remainder of the paper, we adopt the following notation. We refer to the current TEE and computations parameters simply as TEE and computations parameters. We refer to both as parameters. We refer to the constraints sets that delimit the maximum and minimum levels of the parameters as TEE sets and computations sets. Again, we refer to both as parameters sets. There can be multiple or none computations sets per each task, and the same applies to the TEE. Our goal is to complete the plan with the highest possible parameters as the UAV flies and its batteries drain. The term parameters should not be confused with the battery parameters, which we always mention explicitly. These are fixed values that are not replanned by the algorithm.

@@

The remainder of the paper is organized as follows. The overview of dynamic planning is provided in Section~\ref{sec:prob}, along with a suitable model for the position and energy. The algorithm is proposed in Section~\ref{sec:algo}. Section~\ref{sec:experimental} presents the result and showcases the performances. We propose some conclusions in Section~\ref{sec:conclusion}. Finally, we provide some additional information in the Appendix~\ref{app:proof-eqv} and examples in the Appendix~\ref{app:simulator}.

@@

Let us suppose the TEE at stage $i$ can be altered with $\rho$ TEE parameters and the computations with $\sigma$ computations parameters. We then express the TEE as a continuous twice differentiable function $\varphi_i:\mathbb{R}^2\times\mathbb{R}^\rho\rightarrow\mathbb{R}$ of a point and the TEE parameters which returns a metric of the distance between the point and the nominal trajectory. We express and the computations as the value of the computations parameters. We discuss the concrete meaning of the value of TEE parameters in Subsection~\ref{sec:model}), and the of the computations parameters in Subsection~\ref{sec:computations-model}.

Let us further adopt the following mathematical notation.  $[a]$ is the set $\{0,1,\dots,a\}$, $[a]^+$ the set $[a]/\{0\}$, $\langle a_1,a_2,...,a_n\rangle$ an ordered list of $n$ elements, $c^0$ the optimal value of parameter $c$,and $\underline{c},\overline{c}$ its lower and upper bound retrieved from a plan specification. The plan specification is a file which specifies all the stages and triggering points along the plan. We discuss the format in Subsection~\ref{sec:plan}. Moreover, we use the notation $c_{i,j}$ to specify the $j$-th item of the $i$-th parameters set $c_i$.

\begin{defn}[Stage, plan and triggering point]\label{def:mission}
  A \emph{stage} $\Gamma_i$ at time instant $k$ of a plan $\Gamma$ is defined as the ordered list
  \begin{equation}\label{eq:mission}\begin{split}
    \Gamma_i:=\{\langle&\varphi_i(\mathbf{p}_k,c_{i,1},\dots,c_{i,\rho}),c_{i,\rho+1},\dots,c_{i,\rho+\sigma}\rangle\\\mid
    &\,\exists\,\,\mathbf{p}_k,\,\varphi_i(\mathbf{p}_k,c_{i,1},\dots,c_{i,\rho})\in\mathbb{C}_i,\,\\
      &\,\forall j\in[\sigma]^+,\,c_{i,\rho+j}\in\mathbb{S}_{i,j}\,\},
  \end{split}\end{equation}
  where $\mathbb{C}_i:=[\underline{c}_i,\overline{c}_i]\subseteq\mathbb{R}$ is the TEE set, and $\mathbb{S}_{i,j}:=[\underline{c}_{i,\rho+j},\overline{c}_{i,\rho+j}]\subseteq\mathbb{Z}_{\geq 0}$ the $j$-th computation set. $\mathbf{p}_k$ is a point of a UAV flying at an assigned altitude $h\in\mathbb{R}_{>0}$ w.r.t. some inertial navigation frame $\mathcal{O}_W$. We note that while the TEE can differ for each stage, the tasks remain the same. However, the user can inhibit or enable a computation varying its computation set.

  The \emph{plan} is a finite state machine $\Gamma$ where the state-transition function $\bigcup_i{\Gamma_i}\times\mathbb{R}^2\rightarrow\bigcup_i{\Gamma_i}$ maps a stage and a point to the next stage. The point $\mathbf{p}_{\Gamma_{i}}$ that allows the transition between $\Gamma_i$ and $\Gamma_{i+1}$ is called \emph{triggering point}.

@@

  A slice of a finite state machine in Figure~\ref{fig:state-machine} shows the transition between the stages. The triggering point $\mathbf{p}_{i-1}$ allows the transition to stage $\Gamma_i$. The UAV remains in the stage with any generic point $\mathbf{p}_k$. It eventually enters the stage $\Gamma_{i+1}$ with the triggerint point $\mathbf{p}_i$ and so on, until it reaches the last triggering point.

  The algorithm takes as input the plan, the triggering points (using the plan specification) and some initial guesses that are necessary for the position and energy estimation. It outputs the new position, the instantaneous energy consumption, and the control input--an action performed evolving the state.
\end{defn}

@@

We specify the stages $\bigcup_i{\Gamma_i}$, triggering points $\bigcup_i\mathbf{p}_{\Gamma_i}$, and the plan $\Gamma$ using a specific file format--the plan specification. We refer the reader to Appendix~\ref{app:plan-example} for an example.

The first stage $\Gamma_1$ is specified as the first line of the file
\begin{algorithmic}[1]
  \State\textbf{\texttt{$k_e$;$r$;$\varphi_1(\mathbf{p}_k,c_{1,1},\dots,c_{1,\rho})$}}\label{code:stage}
\end{algorithmic}
where $k_e$ is a convergence value, $r$ a rotation, and $\varphi_1$ the TEE. We refer the reader to the next subsection where we discuss the physical meaning of the convergence value and rotation. The rotation is expressed in binary format with zero being the counter-clockwise rotation, one the clockwise rotation.

The next line specifies the triggering point $\mathbf{p}_{\Gamma_1}$
\begin{algorithmic}[1]
  \setcounterref{ALG@line}{code:stage}
  \State\textbf{\texttt{$x$,$y$}}\label{code:trigger-pt}
\end{algorithmic}
where $x$ and $y$ are the two coordinates of the point. The UAV reaches the point within a given radius (a parameter of the algorithm) and enters the next stage $\Gamma_2$. The coordinates can be expressed in function of the TEE parameters $c_1$. Generally, one can express the coordinates in function of any TEE parameters $c_i$, propagating the information therein if necessary (see Figure~\ref{fig:state-machine2}).

@@

We refer the reader to the example in Appendix~\ref{app:plan-example} for a detailed implementation. In the example, the first TEE parameter is propagated to all the following TEEs and triggering points.

The following $\rho+\sigma$ lines
\begin{algorithmic}[1]
  \setcounterref{ALG@line}{code:trigger-pt}
  \State\textbf{\texttt{[$\underline{c}_i$,$\overline{c}_i$]}}\label{code:parameter-set}
\end{algorithmic}
specifies the lower $\underline{c}_i$ and upper $\overline{c}_i$ bound of the $i$-th parameter set. They are in ascending order and the TEE parameters sets are specified before the computations parameters sets.

@@

Let us assume the set
\begin{equation}\label{eq:area}
  \mathcal{P}_i:=\{\mathbf{p}_k\mid\varphi_i(\mathbf{p}_k,c_{i,1},\dots,c_{i,\rho})\in\mathbb{C}_i\},
\end{equation}
delimits the area where the $i$-th TEE $\varphi_i$ is free to evolve using $\rho$ adjustments $\mathbf{c}_i:=c_{i,1},...,c_{i,\rho}$ (the gray area in Figure~\ref{fig:overview}). Phisically, $\varphi_i$ is s function of the two coordinates and is satisfied when a point $\mathbf{p}_k$ is on the TEE. Physically, this means the UAV is flying exactly over the nominal trajectory. The TEE parameters allows to change the TEE. They are a way to alter the nominal trajectory in the initial plan.

In fact, the algorithm uses the set from Equation~(\ref{eq:area}) to alter the optimal TEE parameters $\mathbf{c}_i^0$, so $\varphi_i(\mathbf{p}_k,\mathbf{c}_i^0)$ has the highest instantaneous energy consumption (while still respecting the energy budget). It guides the UAV to the new position $\mathbf{p}_{k+1}$ using the function $\varPhi:=\varphi_i(\mathbf{p}_k,\mathbf{c}_i^0)$, computing its vector field $\nabla\varPhi:=(\partial\varPhi/\partial x,\partial\varPhi/\partial y)$, and deriving the direction to follow in the form of velocity vector
\begin{equation}\label{eq:pd}
  \dot{\mathbf{p}}_d(\mathbf{p}_k):=E\nabla\varPhi-k_e\varPhi\nabla\varPhi,\,\,\,E=\begin{bmatrix}
    0&1\\-1&0
  \end{bmatrix},
\end{equation}
where $E$ specifies the rotation (which influence the tracking direction), and $k_e\in\mathbb{R}_{\geq 0}$ the gain to adjusts the speed of convergence. The direction the velocity vector $\dot{\mathbf{p}}_d$ is pointing at is generally different from the course heading due to the atmospheric interference, such as wind ($w\in\mathbb{R}$ in Figure~\ref{fig:overview}).

% fixed up to here!
% idea; write control as parameters; write B as time-dependent, which multiplied TEE parameters gives energy due to trajectory (approx). By computations due to computations (from powprof).

@@

We model the energy using some energy coefficients $\mathbf{q}\in\mathbb{R}^m$ that characterize the energy signal. We refer to the instantaneous energy consumption evolution simply as energy signal. The coefficients are derived from Fourier analysis (the size of the energy coefficients vector $m$ is related to the order of a Fourier series) and decomposes the evolution in energy due to the path and computations. This approach is adapted from our earlier work on computational energy analysis~\cite{seewald2019coarse, seewald2019component}, and energy estimation of a fixed-wing UAV~\cite{seewald2020mechanical}.

We show a relation between the energy signal and the energy coefficients in Theorem~\ref{thm:state-vs-energy}, but the two are different. We show after the main results how this approach allows us variability in terms of the systems behaving periodically, piece-wise periodically, or merely linearly with sporadic periodicity.

@@

We use extensively the following assumption for the purpose of energy signal modeling.
\begin{assm}[Energy evelution periodicity]\label{assm:periodic} 
Given two time instants $k_1,k_2$ s.t. $k_1>k_2$ and a constant value $n\in\mathbb{Z}_{> 0}$
\begin{equation}
  \abs{y_{k}-y_{k+n}}\in\mathbb{E}\subset\mathbb{R}_{\geq 0} \,\,\,\forall k\in[k_1,k_2].
\end{equation}
\end{assm}

Physically this means that the time evolution of the instantaneous energy consumption is assumed to be approximately periodic.

@@

\begin{lem}[Signals equality]\label{lem:eqv}Suppose there is no state and output uncertainty (the contributions $\mathbf{w}\in\mathbb{R}^m,v\in\mathbb{R}$ are zeros) and the control is zero. Suppose further matrices $A,C$ are described by Equation~(\ref{eq:state-details}), and the initial guess $\mathbf{q}_0$ is $\mathbf{q}_0=\begin{bmatrix}a_0 & a_1/2 & b_1/2 & \cdots & a_r/2 & b_r/2\end{bmatrix}$. Then Equation~(\ref{eq:fourier}) describes the same signal as Equation~(\ref{eq:state-perf}).
\end{lem}
\begin{proof}
The equivalence of the models is trivial and the equality of the two signals is achieved by a proper choice of items of matrices $A,C$ and the initial guess $\mathbf{q}_0$. We refer the reader to Appendix~\ref{app:proof-eqv} for a formal proof of the equality.
\end{proof}

Let us consider the discretized version $A_d:=A\Delta t+I$ of the matrix $A$. We denote the continuous and discrete time with the standard notation $t,k$. We use the forward Euler approximation with a small enough value of the time step $\Delta t$. One can guarantee to have the same outputs rather than an approximation with $A=e^{At}$.

Let us suppose that at time instant $k$ the plan reached $i$-th stage. The control along with the input matrix
\begin{equation}\label{eq:state-control}\begin{split}
  \mathbf{u}&=\begin{bmatrix}c_{i,1} & \cdots & c_{i,\rho} & c_{i,\rho+1} & \cdots & c_{i,\rho+\sigma}\end{bmatrix}^T,\\
  B\mathbf{u}&:=B_i(\mathbf{u}_k-\mathbf{u}_{k-1}),\,\,\,B_i=\left[\begin{array}{c} \nu_i\\\mathbf{0}\end{array}\right],
\end{split}\end{equation}
where $\mathbf{u}\in\mathbb{R}^n$ is the control with $n=\rho+\sigma$, $B\in\mathbb{R}^{m\times n}$, and $\mathbf{u}_{-1}=\mathbf{0}$. Moreover, the top row of $B$ contains gain factors $\nu_i=\begin{bmatrix}\nu_{i,1} & \cdots & \nu_{i,\rho} & \nu_{i,\rho + 1}& \cdots & \nu_{i,\rho+\sigma}\end{bmatrix}$, quantifying the contribution of a given parameter to the instantaneous energy consumption. The other entries of $B$ are zeros.

Equation~(\ref{eq:state-control}) accounts for the energy due to the change of parameters $\mathbf{u}_k-\mathbf{u}_{k-1}$. For instance, when the TEE $\varphi_0$ is a circle (see Figure~\ref{fig:overview}), a decrement in the TEE parameter $c_{0,1}$--the radius of the circle--adds a negative contribution. It thus simulates the lowering of instantaneous energy consumption $\nu_{0,1}(c_{0,1}-c_{0,1}^-)<0$. The first $\rho$ gain factors quantifies the contribution of the $i$-th TEE parameters. A first guess of these parameters is derived empirically. They are later refined by the algorithm. The following $\sigma$ gain factors quantifies the contribution of the $i$-th computations parameters. They are derived using the modeling tool. We use the gain factors to select the optimal control $\mathbf{u}^0$.

\subsection{Energy evolution due to computations}
\label{sec:computations-model}

The energy cost of the computations (the set of tasks being executed on an embedded board on the UAV) is assessed using \powprof{}, an open-source modeling tool presented in our previous work~\cite{seewald2019coarse}. The tool measures the instantaneous computational energy consumption of software components within the QoS sets. It builds an energy model: a linear interpolation, one per each task. In a ROS based system, it requires the user to implement one or more ROS nodes changing the computational load by node-specific ROS parameters. The approach has been tested in simulation in our previous work~\cite{zamanakos2020energy}.

Specifically, the mission $\mathcal{M}$ contains a set of ordered lists with tasks (recall Definition~\ref{def:mission}). These tasks are implemented by software components (such as ROS nodes in a ROS based system) $\Psi(\mathbf{s}_i):=\langle\psi_1(s_{i,1}),\dots,\psi_\sigma(s_{i,\sigma})\rangle$. They input the desired and output the actual computations
\begin{equation}\label{eq:qos-def}
  \mathbf{s}_i:=\left\{\langle s_{i,1},\dots ,s_{i,\sigma}\rangle\mid \psi_j(s_{i,j})\in\mathbb{S}_{i}\,\forall j\in[\sigma]^+\right\},
\end{equation}
where $s_{i,j}:\mathbb{Z}_{\geq 0}\rightarrow\mathbb{Z}_{\geq 0}$ returns the $j$-th desired computation at stage $i$, $\mathbf{s}_i\in\mathbb{S}_i\subseteq\mathbb{Z}_{\geq 0}^\sigma$ the union of all QoS sets given $\mathbb{S}_i:=\bigcup_{j\in[\sigma]^+}{\mathbb{S}_{i,j}}$.

For instance, if the task $\psi_1$ is a CNN object detector, $s_{1,1}$ corresponds to the computation frames-per-second (fps) rate changing the detection frequency. The algorithm outputs the control input which contains the desired fps rate $s_{1,1}\in\mathbb{S}_{1,1}$. The task $\psi_1(s_{1,1})$ outputs the actual fps rate $s_{k,1}$ which might differ from the desired one (e.g., the CNN object detector might not be able to reach a high fps rate with the current computational resources).

>>>>

%\begin{figure*}[t]
%  \centering
%  \input{\figpath/icra-2021_overview2.tikz}
%  \caption{The mission consists of circles and lines covering a polygon. The UAV heads to $\varphi_0$ initially, using the desired velocity vector $\dot{\mathbf{p}}_d$. It later heads to $\varphi_0^-$ reducing the radius $c_{0,1}$ to satisfy the battery constraints. The UAV then converges to $\varphi_1$ in stage $\mathcal{M}_1$, $\varphi_2$ in stage $\mathcal{M}_2$, and so on (the circle $\varphi_2$ is not visible in the figure; it connects $\varphi_1$ and $\varphi_3$).}
%  \label{fig:overview}
%\end{figure*}

@@

Many scenarios involving unmanned aerial vehicles (UAVs), such as precision agriculture, search and rescue, and surveillance, require high autonomy but have limited energy budgets. A typical example of these scenarios is a UAV flying a path and performing some on-board computational tasks. For instance, the UAV might detect ground patterns and inform other ground-based actors with little human interaction. We refer to the computational tasks that can be dynamically replanned as computations. We are interested in energy-aware dynamic planning, which would find optimal tradeoffs between the path, computations, and energy requirements. Current generic planning solutions for outdoor UAVs do not plan path and computations dynamically, nor are energy-aware. They are often semi-autonomous; the plan--the path and computations--is static and usually defined using planning software~\cite{daponte2019review}. Such a state of practice has prompted us to propose an \emph{energy-aware dynamic planning algorithm} for UAVs. The algorithm combines and generalizes some of the past body of knowledge on mobile robot planning problems and addresses the increasing \emph{computational demands} and their relation to energy consumption, path, and autonomy for the UAV planning problem.

@@

Unlike most of the past planning algorithms literature, our algorithm plans both the path and computations. We guide the UAV using a vector field~\cite{de2017guidance} that converges smoothly to a planned trajectory. The use of vector fields for planning is widely discussed in the literature~\cite{lindemann2005smoothly,gonccalves2010vector,panagou2014motion,zhou2014vector,kapitanyuk2017guiding}. We denote the trajectory with a mathematical function that we refer to as the trajectory explicit equation (TEE). The path is composed of multiple TEEs. We express the computations with a modeling tool presented in our previous work~\cite{seewald2019coarse}. The algorithm alters the TEE and computations within some parameters using output model predictive control (MPC). These parameters are the maximum, minimum, and current levels of the TEE and computations. The user specifies the parameters in an initial plan with an energy budget. The energy budget is the battery capacity and some other battery parameters. In the remainder of the paper, we adopt the following notation. We refer to the current TEE and computations parameters simply as TEE and computations parameters. We refer to both as parameters. We refer to the constraints sets that delimit the maximum and minimum levels of the parameters as TEE sets and computations sets, and to both as parameters sets. There can be multiple or none computations sets per each task, and the same applies to the TEE. Our goal is to complete the plan with the highest possible parameters as the UAV flies and its batteries drain. The term parameters should not be confused with the battery parameters, which we always mention explicitly. These are fixed values that are not replanned by the algorithm.

@@

The remainder of the paper is organized as follows. The overview of dynamic planning is provided in Section~\ref{sec:prob}, along with a suitable model for the position and energy. The algorithm is proposed in Section~\ref{sec:algo}. Section~\ref{sec:experimental} presents the result and showcases the performances. We propose some conclusions in Section~\ref{sec:conclusion}. Finally, we provide some additional information and examples in the Appendixes~\ref{app:proof-eqv}--\ref{app:simulator}.

@@

Let us further assume the TEE at stage $i$ can be altered with $\rho$ TEE parameters and the computations with $\sigma$ computations parameters. We then express the TEE as a continuous twice differentiable function $\varphi_i:\mathbb{R}^2\times\mathbb{R}^\rho\rightarrow\mathbb{R}$ of a point and the TEE parameters which returns a metric of the distance between the point and the nominal trajectory. We express the computations as the value of the computations parameters. We discuss the concrete meaning of the value of TEE parameters in Subsection~\ref{sec:model}), and of the computations parameters in Subsection~\ref{sec:computations-model}.

Let us adopt the following mathematical notation. Given an integer $a$, $[a]$ is the set $\{0,1,\dots,a\}$, $[a]^+$ the set $[a]/\{0\}$. $c_{i,j}$ specifues the $j$-th parameter of the $i$-th parameters set $c_i$. $c_{i,j}^0$ is the optimal value of the parameter $c_{i,j}$, and $\underline{c_{i,j}},\overline{c_{i,j}}$ its lower and upper bounds. We store these bounds in the plan specification, which format we discuss in Subsection~\ref{sec:plan}. The set $\langle c_{i,1},c_{i,2},...,c_{i,n}\rangle$ is an ordered list of $n$ parameters.

\begin{defn}[Stage, plan and triggering point]\label{def:mission}
  A \emph{stage} $\Gamma_i$ at time instant $k$ of a plan $\Gamma$ is defined as the ordered list
  \begin{equation}\label{eq:mission}\begin{split}
    \Gamma_i:=\{\langle&\varphi_i(\mathbf{p}_k,c_{i,1},\dots,c_{i,\rho}),c_{i,\rho+1},\dots,c_{i,\rho+\sigma}\rangle\\\mid
    &\,\exists\,\,\mathbf{p}_k,\,\varphi_i(\mathbf{p}_k,c_{i,1},\dots,c_{i,\rho})\in\mathbb{C}_i,\,\\
      &\,\forall j\in[\sigma]^+,\,c_{i,\rho+j}\in\mathbb{S}_{i,j}\,\},
  \end{split}\end{equation}
  where $\mathbb{C}_i:=[\underline{c}_i,\overline{c}_i]\subseteq\mathbb{R}$ is the TEE set, and $\mathbb{S}_{i,j}:=[\underline{c}_{i,\rho+j},\overline{c}_{i,\rho+j}]\subseteq\mathbb{Z}_{\geq 0}$ the $j$-th computation set. $\mathbf{p}_k$ is a point of a UAV flying at an assigned altitude $h\in\mathbb{R}_{>0}$ w.r.t. some inertial navigation frame $\mathcal{O}_W$.

  The \emph{plan} is a finite state machine $\Gamma$ where the state-transition function $\bigcup_i{\Gamma_i}\times\mathbb{R}^2\rightarrow\bigcup_i{\Gamma_i}$ maps a stage and a point to the next stage. The point $\mathbf{p}_{\Gamma_{i}}$ that allows the transition between $\Gamma_i$ and $\Gamma_{i+1}$ is called \emph{triggering point}.
\end{defn}

@@

A slice of a finite state machine in Figure~\ref{fig:state-machine} shows the transition between the stages. The triggering point $\mathbf{p}_{i-1}$ allows the transition to stage $\Gamma_i$. The UAV remains in the stage with any generic point $\mathbf{p}_k$. It eventually enters the stage $\Gamma_{i+1}$ with the triggerint point $\mathbf{p}_{\Gamma_i}$ and so on, until it reaches the last triggering point.

The algorithm takes as input the plan and some initial guesses that are necessary for the position and energy estimation. It outputs the new position, the instantaneous energy consumption, and the control input--an action performed evolving the state.

@@

The algorithm reads the plan $\Gamma$ using a specific file format--the plan specification. We refer the reader to Appendix~\ref{app:plan-example} for an example.

The first stage $\Gamma_1$ is expressed in the first line of the file
\begin{algorithmic}[1]
  \State\textbf{\texttt{$k_e$;$r$;$\varphi_1(\mathbf{p}_k,c_{1,1},\dots,c_{1,\rho})$}}\label{code:stage}
\end{algorithmic}
where $k_e$ is a convergence value, $r$ a rotation, and $\varphi_1$ the TEE. We refer the reader to the next subsection where we discuss the physical meaning of the convergence value and rotation. The rotation is expressed in binary format with zero being the counter-clockwise rotation, one the clockwise rotation.

The next line specifies the triggering point $\mathbf{p}_{\Gamma_1}$
\begin{algorithmic}[1]
  \setcounterref{ALG@line}{code:stage}
  \State\textbf{\texttt{$x$,$y$}}\label{code:trigger-pt}
\end{algorithmic}
where $x$ and $y$ are the two coordinates of the point. The UAV reaches the point within a given radius (a parameter of the algorithm) and enters the next stage $\Gamma_2$. The coordinates can be expressed in function of the TEE parameters $c_{1,1},\dots,c_{1,\rho}$. Generally, one can express the coordinates in function of the $i$-th TEE parameters $c_{i,1},\dots,c_{i,\rho}$, or any previous TEE parameters, propagating the information therein if necessary. See Figure~\ref{fig:state-machine2} where we neglected the $\sigma=0$ computations parameters for simplicity $c_i=c_{i,1},\dots,c_{i,\rho}$.

@@

We refer the reader again to the example in Appendix~\ref{app:plan-example} for a detailed implementation. In the example, the first TEE parameter is propagated to all the following TEEs and triggering points.

The following $\rho+\sigma$ lines
\begin{algorithmic}[1]
  \setcounterref{ALG@line}{code:trigger-pt}
  \State\textbf{\texttt{[$\underline{c}_{1,j}$,$\overline{c}_{1,j}$]}}\label{code:parameter-set}
\end{algorithmic}
specifies the lower $\underline{c}_{1,j}$ and upper $\overline{c}_{1,j}$ bound of the $j$-th parameter set. They are in ascending order and the TEE parameters sets are specified before the computations parameters sets.

@@

Let us assume that the set
\begin{equation}\label{eq:area}
  \mathcal{P}_i:=\{\mathbf{p}_k\mid\varphi_i(\mathbf{p}_k,c_{i,1},\dots,c_{i,\rho})\in\mathbb{C}_i\},
\end{equation}
delimits the area where the $i$-th TEE $\varphi_i$ is free to evolve using the TEE parameters $c_{i,1},...,c_{i,\rho}$ (the gray area in Figure~\ref{fig:overview}). $\varphi_i$ is a function of the two coordinates and is equal to zero when a point $\mathbf{p}_k$ is on the TEE. Physically, this means the UAV is flying exactly over the nominal trajectory. The TEE parameters allows to change the TEE. They are a way to alter the nominal trajectory in the initial plan.
In fact, the algorithm uses the set from Equation~(\ref{eq:area}) to alter the optimal TEE parameters $c_i^0$ (here we again neglect the $\sigma=0$ computations parameters for simplicity), such that $\varphi_i(\mathbf{p}_k,c_i^0)$ has the highest instantaneous energy consumption (while still respecting the energy budget).

We derive the new position $\mathbf{p}_{k+1}$ using the function $\varPhi:=\varphi_i(\mathbf{p}_k,\mathbf{c}_i^0)$, computing its vector field $\nabla\varPhi:=(\partial\varPhi/\partial x,\partial\varPhi/\partial y)$, and the direction to follow in the form of velocity vector
\begin{equation}\label{eq:pd}
  \dot{\mathbf{p}}_d(\mathbf{p}_k):=E\nabla\varPhi-k_e\varPhi\nabla\varPhi,\,\,\,E=\begin{bmatrix}
    0&1\\-1&0
  \end{bmatrix},
\end{equation}
where $E$ specifies the rotation (which influence the tracking direction), and $k_e\in\mathbb{R}_{\geq 0}$ the gain to adjusts the speed of convergence. The direction the velocity vector $\dot{\mathbf{p}}_d$ is pointing at is generally different from the course heading due to the atmospheric interference, such as wind ($w\in\mathbb{R}$ in Figure~\ref{fig:overview}).

@@

We model the energy using some energy coefficients $\mathbf{q}\in\mathbb{R}^m$ that characterize the energy signal. We refer to the instantaneous energy consumption evolution simply as energy signal. The coefficients are derived from Fourier analysis (the size of the energy coefficients vector $m$ is related to the order of a Fourier series). Here we derive the energy model. Later we assess the energy contribution of the computations in Subsection~\ref{sec:computations-model}. This approach is adapted from our earlier work on computational energy analysis~\cite{seewald2019coarse, seewald2019component}, and energy estimation of a fixed-wing UAV~\cite{seewald2020mechanical}.

We show a relation between the energy signal and the energy coefficients in Theorem~\ref{thm:state-vs-energy}. We show after the main results how this approach allows us variability in terms of the systems behaving periodically, piece-wise periodically, or merely linearly with sporadic periodicity.

@@

We use extensively the following assumption for the purpose of energy signal modeling.
\begin{assm}[Energy signal periodicity]\label{assm:periodic} 
Given two time instants $k_1,k_2$ s.t. $k_1>k_2$ and a constant value $n\in\mathbb{Z}_{> 0}$
\begin{equation}
  \abs{y_{k}-y_{k+n}}\in\mathbb{E}\subset\mathbb{R}_{\geq 0} \,\,\,\forall k\in[k_1,k_2].
\end{equation}
\end{assm}

Physically this means that the energy signal is assumed to be approximately periodic.

@@

\begin{lem}[Signals equality]\label{lem:eqv}Suppose there is no state and output uncertainty (the contributions $\mathbf{w}\in\mathbb{R}^m,v\in\mathbb{R}$ are zeros) and the control is zero. Suppose further matrices $A,C$ are described by Equation~(\ref{eq:state-details}), and the initial guess $\mathbf{q}_0$ is $\mathbf{q}_0=\begin{bmatrix}a_0 & a_1/2 & b_1/2 & \cdots & a_r/2 & b_r/2\end{bmatrix}$. Then Equation~(\ref{eq:fourier}) describes the same signal as Equation~(\ref{eq:state-perf}).
\end{lem}
\begin{proof}
The equivalence of the models is trivial and the equality of the two signals is achieved by a proper choice of items of matrices $A,C$ and the initial guess $\mathbf{q}_0$. We refer the reader to Appendix~\ref{app:proof-eqv} for a formal proof of the equality where we justify the choices of the items of the matrices and of the initial guess. 
\end{proof}

Let us consider the discretized version $A_d:=A\Delta t+I$ of the matrix $A$. We denote the continuous and discrete time with the standard notation $t,k$. We use the forward Euler approximation with a small enough value of the time step $\Delta t$. We note that one can guarantee to have the same outputs rather than an approximation with $A_d:=e^{At}$.

Let us suppose that at time instant $k$ the plan reached $i$-th stage $\Gamma_i$. The control along with the input matrix
\begin{equation}\label{eq:state-control}\begin{split}
  \mathbf{u}&=\begin{bmatrix}c_{i,1} & \cdots & c_{i,\rho} & c_{i,\rho+1} & \cdots & c_{i,\rho+\sigma}\end{bmatrix}^T,\\
  B\mathbf{u}&:=B_i(\mathbf{u}_k-\mathbf{u}_{k-1}),\,\,\,B_i=\left[\begin{array}{c} \nu_i\\\mathbf{0}\end{array}\right],
\end{split}\end{equation}
where $\mathbf{u}\in\mathbb{R}^n$ is the control with $n=\rho+\sigma$, $B\in\mathbb{R}^{m\times n}$, and $\mathbf{u}_{-1}=\mathbf{0}$. Moreover, the top row of $B$ contains gain factors $\nu_i=\begin{bmatrix}\nu_{i,1} & \cdots & \nu_{i,\rho} & \nu_{i,\rho + 1}& \cdots & \nu_{i,\rho+\sigma}\end{bmatrix}$, quantifying the contribution of a given parameter to the instantaneous energy consumption. The other entries of $B$ are zeros.

The first $\rho$ gain factors quantifies the contribution of the $i$-th TEE parameters. A first guess of these parameters is derived empirically. They are later refined by the algorithm. The following $\sigma$ gain factors quantifies the contribution of the $i$-th computations parameters. They are derived using the modeling tool. We use the gain factors to select the optimal control $\mathbf{u}^0$.

Equation~(\ref{eq:state-control}) accounts for the energy due to the change of parameters $\mathbf{u}_k-\mathbf{u}_{k-1}$. For instance, when the TEE $\varphi_0$ is a circle (see Figure~\ref{fig:overview}), a decrement in the TEE parameter $c_{0,1}$--the radius of the circle--adds a negative contribution. It thus simulates the lowering of instantaneous energy consumption $\nu_{0,1}(c_{0,1}-c_{0,1}^-)<0$.

\subsection{Computations parameters energy contribution}
\label{sec:computations-model}

Let us recall from Definition~\ref{def:mission} that the $i$-th stage $\Gamma_i$ of the plan $\Gamma$ contains the computations parameters which characterize the computations. We assess the energy cost of these computations using \powprof{}, an open-source modeling tool presented in our previous work~\cite{seewald2019coarse}. We assume the UAV carries an embedded board that runs the computations. The tool measures the instantaneous energy consumption of a subset of possible computations parameters within the computations sets and builds an energy model: a linear interpolation, one per each computation.

Specifically, the computations are implemented by software components, such as ROS nodes in a ROS based system. The user implements these nodes such that they change the computational load by node-specific ROS parameters--the computations parameters. For instance, if the computation is a CNN object detector, the computation parameter $c_{i,\rho+1}$ corresponds frames-per-second (fps) rate and it changes the detection frequency. We note that while the TEE can differ for each stage, the tasks remain the same. However, the user can inhibit or enable a computation varying its computation set.

Let us further define $g:\mathbb{Z}_{\geq 0}^{\sigma}\rightarrow\mathbb{R}_{\geq 0}$ as the instantaneous computational energy consumption value obtained using \powprof{}
\begin{equation}\label{eq:energy-comp}
  \begin{bmatrix}\nu_{i,\rho+1} \\ \vdots \\ \nu_{i,\rho+\sigma}\end{bmatrix}^T=\begin{bmatrix}g(c_{i,\rho+1}) \\ \vdots \\ g(c_{i,\rho+\sigma})\end{bmatrix}/\begin{bmatrix}c_{i,\rho+1} & \cdots & c_{i,\rho+\sigma}\end{bmatrix}.
\end{equation}
Moreover, let $g(\{\emptyset\})$ be zero.

\leavevmode\thispagestyle{empty}\newpage
\leavevmode\thispagestyle{empty}\newpage
